---
title: "Day 3 R"
author: "Alicia Burris"
date: "9/19/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##load 2 data frames
```{r}
generation <- read.csv(here::here("data/generation.csv"), 
                       stringsAsFactors = F)
imports <- read.csv(here::here("data/imports.csv"), 
                    stringsAsFactors = F)
```
##change date time variables to datetime class
```{r}
library(lubridate)
generation$datetime <- as_datetime(generation$datetime)
class(generation$datetime)
head(generation$datetime)
imports$datetime <- as_datetime(imports$datetime)
class(imports$datetime)
head(imports$datetime)
```
##restructure generation to go from wide data to long data
```{r}
head(generation)
library(reshape2)
long_gen <- melt(generation, id.vars = "datetime",
                 variable.name = "source",
                 value.name = "usage")
head(long_gen)
```
##merge imports and generation dataframes by shared column
```{r}
merged_energy <- merge(generation, imports, by = "datetime")
```
#melt previously merged data frame to go from wide to long data
```{r}
long_merged_energy <- melt(merged_energy, id.vars = "datetime",
                 variable.name = "source",
                 value.name = "usage")
```

##dplyr
```{r}
library(dplyr)
tmp <- select(merged_energy, biogas, biomass, geothermal, solar)
names(tmp)
```

##helpers
```{r}
tmp <- select(merged_energy, contains("hydro"), starts_with("bio"))
names(tmp)
```

##filter
```{r}
tmp <- filter(merged_energy, imports > 7000, natural_gas < 7000)
nrow(tmp)
head(tmp)
```

##mutate
```{r}
tmp <- mutate(long_merged_energy, log_usage = log(usage))
head(tmp)
```

##multiple columns
```{r}
tmp <- mutate(long_merged_energy, log_usage = log(usage), usage2 = usage^2, usage3 = usage^3)
head(tmp)
```

##summarize
```{r}
summarize(long_merged_energy, total = sum(usage, na.rm = T))
summarize(long_merged_energy, mean_cons = mean(usage, na.rm = T))
```

##pipes
```{r}
long_merged_energy %>%
  filter(source == "geothermal") %>%
  select(-datetime) %>%
  mutate(log_usage = log(usage)) %>%
  summarize(mean_log_usage = mean(log_usage, na.rm = T))
```

##another example
```{r}
merged_energy %>%
  select(-datetime) %>%
  mutate(total_usage = rowSums(., na.rm = T)) %>%
  summarize(total_usage = sum(total_usage, na.rm = T))
```
```{r}
merged_energy %>% 
  select(-datetime) %>% 
  mutate(total_usage =  rowSums(., na.rm = T)) %>% 
  summarize(total_usage = sum(total_usage, na.rm = T))
```

##try piping
```{r}
merged_energy %>% 
  select(contains("hydro")) %>% 
  mutate(total_hydro = rowSums(.,na.rm = T)) %>% 
  summarize(mean_hydro = mean(total_hydro, na.rm = T))
```
##group_by with summarize example
```{r}
long_merged_energy %>% 
  group_by(source) %>%
  summarize(sum_usage = sum(usage, na.rm = T))
```

##group by exercise
```{r}
merged_energy %>% 
  select(datetime, contains("hydro"), contains("bio")) %>% 
  melt(id.vars = "datetime",
       variable.name = "source",
       value.name = "usage") %>% 
  group_by(source) %>% 
  summarize(mean_usage = mean(usage, na.rm = T))
```

















